---
title: "Reading the Odds"
subtitle: "Multivariable Logistic Regression --- from y = mx + b to Log-Odds"
author:
  - name: "John Paul Bisciotti"
    affiliations: "Lead Data Scientist, Department of Defense" 
format:
  revealjs:
    theme: [default, custom.scss]
    slide-number: false
    menu: false
    width: 1280
    height: 720
title-slide-attributes:
  data-background-color: "#1e3b33"
execute:
  echo: false
  warning: false
---

```{r}
#| label: setup
#| include: false
source("plots.R")
source("notes.R")
```

```{r}
#| label: sigmoid-figure
#| include: false
# Generate the light-blue sigmoid with dashed asymptotes once. The image is
# placed on the title slide by title_slide_decorations() at the end of the deck
# (kept out of the preamble so no blank slide is created before Roadmap).

fig_slide_01 <- plot_fig_slide_01()

ggplot2::ggsave(
  "sigmoid.png",
  fig_slide_01,
  width = 4, height = 1.5, dpi = 200,
  bg = "transparent"
)
```

## Disclaimer

The views and opinions expressed in this document are those of the author(s) and do not represent the official position, policy, or endorsement of the Department of Defense (also known as the Department of War), the United States Government, or any of their agencies or components, unless another official document expressly designates them as such. 

```{r}
#| label: title-slide-decorations
#| output: asis
# Decorates the auto title slide (sigmoid image + speaker note). Lives here,
# inside a real slide, rather than in the preamble.
title_slide_decorations("slide-01")
```

## Roadmap

::: {.section-label}
Section 1:  build our understanding of the model, in four steps
:::

:::: {.columns .roadmap-row}

::: {.column width="100%"}

:::

::: {.column .section-two-col width="100%"}

:::

::::

```{r}
#| label: notes-slide-02
#| output: asis
speaker_note("slide-02")
```

## Roadmap

::: {.section-label}
Section 1:  build our understanding of the model, in four steps
:::

:::: {.columns .roadmap-row}

::: {.column width="100%"}
```{r}
#| label: roadmap-panels-1
#| fig-width: 12
#| fig-height: 4
#| fig-dpi: 200
plot_fig_slide_02(1)
```
:::

::: {.column .section-two-col width="100%"}

:::

::::

```{r}
#| label: notes-slide-02-1
#| output: asis
speaker_note("slide-02")
```

## Roadmap

::: {.section-label}
Section 1:  build our understanding of the model, in four steps
:::

:::: {.columns .roadmap-row}

::: {.column width="100%"}
```{r}
#| label: roadmap-panels-2
#| fig-width: 12
#| fig-height: 4
#| fig-dpi: 200
plot_fig_slide_02(2)
```
:::

::: {.column .section-two-col width="100%"}

:::

::::

```{r}
#| label: notes-slide-02-2
#| output: asis
speaker_note("slide-02")
```

## Roadmap

::: {.section-label}
Section 1:  build our understanding of the model, in four steps
:::

:::: {.columns .roadmap-row}

::: {.column width="100%"}
```{r}
#| label: roadmap-panels-3
#| fig-width: 12
#| fig-height: 4
#| fig-dpi: 200
plot_fig_slide_02(3)
```
:::

::: {.column .section-two-col width="100%"}

:::

::::

```{r}
#| label: notes-slide-02-3
#| output: asis
speaker_note("slide-02")
```

## Roadmap

::: {.section-label}
Section 1:  build our understanding of the model, in four steps
:::

:::: {.columns .roadmap-row}

::: {.column width="100%"}
```{r}
#| label: roadmap-panels-4
#| fig-width: 12
#| fig-height: 4
#| fig-dpi: 200
plot_fig_slide_02(4)
```
:::

::: {.column .section-two-col width="100%"}

:::

::::

```{r}
#| label: notes-slide-02-4
#| output: asis
speaker_note("slide-02")
```

## Roadmap

::: {.section-label}
Section 1:  build our understanding of the model, in four steps
:::

:::: {.columns .roadmap-row}

::: {.column width="100%"}
```{r}
#| label: roadmap-panels-5
#| fig-width: 12
#| fig-height: 4
#| fig-dpi: 200
plot_fig_slide_02(4)
```
:::

::: {.column .section-two-col width="100%"}
[Section 2:]{.section-two-label}<br>
interpret a published figure
:::

::::

```{r}
#| label: notes-slide-02-5
#| output: asis
speaker_note("slide-02")
```

## Exact relationship 

```{r}
#| label: glasses-of-water-1
plot_fig_slide_03(1)
```

```{r}
#| label: notes-slide-03-1
#| output: asis
speaker_note("slide-03")
```

## Exact relationship 

```{r}
#| label: glasses-of-water-2
plot_fig_slide_03(2)
```

```{r}
#| label: notes-slide-03-2
#| output: asis
speaker_note("slide-03")
```

## Exact relationship 

```{r}
#| label: glasses-of-water-3
plot_fig_slide_03(3)
```

```{r}
#| label: notes-slide-03-3
#| output: asis
speaker_note("slide-03")
```

## Exact relationship --- C and F

```{r}
#| label: temperature-plot-1
plot_fig_slide_04(1)
```

```{r}
#| label: notes-slide-04-1
#| output: asis
speaker_note("slide-04")
```

## Exact relationship --- C and F

```{r}
#| label: temperature-plot-2
plot_fig_slide_04(2)
```

```{r}
#| label: notes-slide-04-2
#| output: asis
speaker_note("slide-04")
```

## Exact relationship --- C and F

```{r}
#| label: temperature-plot-3
plot_fig_slide_04(3)
```

```{r}
#| label: notes-slide-04-3
#| output: asis
speaker_note("slide-04")
```

## Exact relationship --- C and F

```{r}
#| label: temperature-plot-4
plot_fig_slide_04(4)
```

```{r}
#| label: notes-slide-04-4
#| output: asis
speaker_note("slide-04")
```

## Uncertainty 

```{r}
#| label: balance-scales-1
#| output: asis
plot_fig_slide_05(1)
```

```{r}
#| label: notes-slide-05-1
#| output: asis
speaker_note("slide-05")
```

## Uncertainty 

```{r}
#| label: balance-scales-2
#| output: asis
plot_fig_slide_05(2)
```

```{r}
#| label: notes-slide-05-2
#| output: asis
speaker_note("slide-05")
```

## Uncertainty 

```{r}
#| label: balance-scales-3
#| output: asis
plot_fig_slide_05(3)
```

```{r}
#| label: notes-slide-05-3
#| output: asis
speaker_note("slide-05")
```

## Uncertainty 

```{r}
#| label: balance-scales-4
#| output: asis
plot_fig_slide_05(4)
```

```{r}
#| label: notes-slide-05-4
#| output: asis
speaker_note("slide-05")
```

## Uncertainty --- height and weight

```{r}
#| label: height-and-weight-plot-1
plot_fig_slide_06(1)
```

```{r}
#| label: notes-slide-06-1
#| output: asis
speaker_note("slide-06")
```

## Uncertainty --- height and weight

```{r}
#| label: height-and-weight-plot-2
plot_fig_slide_06(2)
```

```{r}
#| label: notes-slide-06-2
#| output: asis
speaker_note("slide-06")
```

## Uncertainty --- height and weight

```{r}
#| label: height-and-weight-plot-3
plot_fig_slide_06(3)
```

```{r}
#| label: notes-slide-06-3
#| output: asis
speaker_note("slide-06")
```

## Uncertainty --- height and weight

```{r}
#| label: height-and-weight-plot-4
plot_fig_slide_06(4)
```

```{r}
#| label: notes-slide-06-4
#| output: asis
speaker_note("slide-06")
```

## Uncertainty --- height and weight

```{r}
#| label: height-and-weight-plot-5
plot_fig_slide_06(5)
```

```{r}
#| label: notes-slide-06-5
#| output: asis
speaker_note("slide-06")
```

## Uncertainty --- height and weight

```{r}
#| label: height-and-weight-plot-6
plot_fig_slide_06(6)
```

```{r}
#| label: notes-slide-06-6
#| output: asis
speaker_note("slide-06")
```

## Yes-or-no outcome

```{r}
#| label: death-status-1
plot_fig_slide_07(1)
```

```{r}
#| label: notes-slide-07-1
#| output: asis
speaker_note("slide-07")
```

## Yes-or-no outcome

```{r}
#| label: death-status-2
plot_fig_slide_07(2)
```

```{r}
#| label: notes-slide-07-2
#| output: asis
speaker_note("slide-07")
```

## Yes-or-no outcome

```{r}
#| label: death-status-3
plot_fig_slide_07(3)
```

```{r}
#| label: notes-slide-07-3
#| output: asis
speaker_note("slide-07")
```

## Yes-or-no outcome

```{r}
#| label: death-status-4
plot_fig_slide_07(4)
```

```{r}
#| label: notes-slide-07-4
#| output: asis
speaker_note("slide-07")
```

## Yes-or-no outcome

```{r}
#| label: death-status-5
plot_fig_slide_07(5)
```

```{r}
#| label: notes-slide-07-5
#| output: asis
speaker_note("slide-07")
```

## Yes-or-no outcome

```{r}
#| label: death-status-6
plot_fig_slide_07(6)
```

```{r}
#| label: notes-slide-07-6
#| output: asis
speaker_note("slide-07")
```

## Yes-or-no outcome

```{r}
#| label: death-status-7
plot_fig_slide_07(7)
```

```{r}
#| label: notes-slide-07-7
#| output: asis
speaker_note("slide-07")
```

## Odds

```{r}
#| label: ten-people-1
plot_fig_slide_08(1)
```

```{r}
#| label: notes-slide-08-1
#| output: asis
speaker_note("slide-08")
```

## Odds

```{r}
#| label: ten-people-2
plot_fig_slide_08(2)
```

```{r}
#| label: notes-slide-08-2
#| output: asis
speaker_note("slide-08")
```

## Odds

```{r}
#| label: ten-people-3
plot_fig_slide_08(3)
```

```{r}
#| label: notes-slide-08-3
#| output: asis
speaker_note("slide-08")
```

## ln(odds)

```{r}
#| label: number-line-1
plot_fig_slide_09(1)
```

```{r}
#| label: notes-slide-09-1
#| output: asis
speaker_note("slide-09")
```

## ln(odds)

```{r}
#| label: number-line-2
plot_fig_slide_09(2)
```

```{r}
#| label: notes-slide-09-2
#| output: asis
speaker_note("slide-09")
```

## ln(odds)

```{r}
#| label: number-line-3
plot_fig_slide_09(3)
```

```{r}
#| label: notes-slide-09-3
#| output: asis
speaker_note("slide-09")
```

## ln(odds)

```{r}
#| label: number-line-4
plot_fig_slide_09(4)
```

```{r}
#| label: notes-slide-09-4
#| output: asis
speaker_note("slide-09")
```

## ln(odds)

```{r}
#| label: number-line-5
plot_fig_slide_09(5)
```

```{r}
#| label: notes-slide-09-5
#| output: asis
speaker_note("slide-09")
```

## ln(odds)

```{r}
#| label: number-line-6
plot_fig_slide_09(6)
```

```{r}
#| label: notes-slide-09-6
#| output: asis
speaker_note("slide-09")
```

## ln(odds)

```{r}
#| label: number-line-7
plot_fig_slide_09(7)
```

```{r}
#| label: notes-slide-09-7
#| output: asis
speaker_note("slide-09")
```

## ln(odds)

```{r}
#| label: number-line-8
plot_fig_slide_09(8)
```

```{r}
#| label: notes-slide-09-8
#| output: asis
speaker_note("slide-09")
```

## ln(odds)

```{r}
#| label: number-line-9
plot_fig_slide_09(9)
```

```{r}
#| label: notes-slide-09-9
#| output: asis
speaker_note("slide-09")
```

## Logistic regression --- one model, two views

```{r}
#| label: two-views-1
plot_fig_slide_10(1)
```

```{r}
#| label: notes-slide-10-1
#| output: asis
speaker_note("slide-10")
```

## Logistic regression --- one model, two views

```{r}
#| label: two-views-2
plot_fig_slide_10(2)
```

```{r}
#| label: notes-slide-10-2
#| output: asis
speaker_note("slide-10")
```

## Reading a coefficient 

```{r}
#| label: read-coef
plot_fig_slide_11()
```

```{r}
#| label: notes-slide-11
#| output: asis
speaker_note("slide-11")
```

## Odds Ratio

```{r}
#| label: odds-ratio-1
plot_fig_slide_12(1)
```

```{r}
#| label: notes-slide-12-1
#| output: asis
speaker_note("slide-12")
```

## Odds Ratio

```{r}
#| label: odds-ratio-2
plot_fig_slide_12(2)
```

```{r}
#| label: notes-slide-12-2
#| output: asis
speaker_note("slide-12")
```

## Odds Ratio

```{r}
#| label: odds-ratio-3
plot_fig_slide_12(3)
```

```{r}
#| label: notes-slide-12-3
#| output: asis
speaker_note("slide-12")
```

## Odds Ratio

```{r}
#| label: odds-ratio-4
plot_fig_slide_12(4)
```

```{r}
#| label: notes-slide-12-4
#| output: asis
speaker_note("slide-12")
```

## Adjustment

```{r}
#| label: adjustment-2
plot_fig_slide_13(2)
```

```{r}
#| label: notes-slide-13-2
#| output: asis
speaker_note("slide-13")
```

## Adjustment

```{r}
#| label: adjustment-3
plot_fig_slide_13(3)
```

```{r}
#| label: notes-slide-13-3
#| output: asis
speaker_note("slide-13")
```

## Adjustment

```{r}
#| label: adjustment-4
plot_fig_slide_13(4)
```

```{r}
#| label: notes-slide-13-4
#| output: asis
speaker_note("slide-13")
```

## Adjustment

```{r}
#| label: adjustment-5
plot_fig_slide_13(5)
```

```{r}
#| label: notes-slide-13-5
#| output: asis
speaker_note("slide-13")
```

## Adjustment

```{r}
#| label: adjustment-6
plot_fig_slide_13(6)
```

```{r}
#| label: notes-slide-13-6
#| output: asis
speaker_note("slide-13")
```

## Adjustment

```{r}
#| label: adjustment-7
plot_fig_slide_13(7)
```

```{r}
#| label: notes-slide-13-7
#| output: asis
speaker_note("slide-13")
```

## Adjustment

```{r}
#| label: adjustment-8
plot_fig_slide_13(8)
```

```{r}
#| label: notes-slide-13-8
#| output: asis
speaker_note("slide-13")
```

## Adjustment

```{r}
#| label: adjustment-9
plot_fig_slide_13(9)
```

```{r}
#| label: notes-slide-13-9
#| output: asis
speaker_note("slide-13")
```

## Adjustment

```{r}
#| label: adjustment-10
plot_fig_slide_13(10)
```

```{r}
#| label: notes-slide-13-10
#| output: asis
speaker_note("slide-13")
```

## Adjustment

```{r}
#| label: adjustment-11
plot_fig_slide_13(11)
```

```{r}
#| label: notes-slide-13-11
#| output: asis
speaker_note("slide-13")
```

## Adjustment

```{r}
#| label: adjustment-12
plot_fig_slide_13(12)
```

```{r}
#| label: notes-slide-13-12
#| output: asis
speaker_note("slide-13")
```

## &nbsp; {.summary data-background-color="#1e3b33"}

```{=html}
<style>
.reveal .summary h2 { display: none; }                 /* hide the empty title */
.reveal .summary .s14-body {
  position: absolute; top: 27%; left: 6%; right: 6%;
  color: #ffffff; font-size: 1.15em; line-height: 1.5;
}
.reveal .summary .s14-hl { color: #85e7d4; font-weight: 700; }
.reveal .summary .s14-foot {
  position: absolute; bottom: 7%; left: 6.5%;
  color: #cddcf9; font-size: 0.55em;
  font-family: system-ui, "Helvetica Neue", Arial, sans-serif;
}
</style>
```

::: {.s14-body}
Logistic regression models the [log-odds]{.s14-hl} of a yes/no outcome as
a [straight line]{.s14-hl} in one or more predictors; each [e^β^]{.s14-hl} is an
[odds ratio]{.s14-hl} --- the factor by which the odds multiply for a one-unit
increase, holding the other predictors constant; and an odds
ratio of [1]{.s14-hl} is the [anchor]{.s14-hl}: no difference.
:::

[Section 1 complete.  Next: reading a real figure.]{.s14-foot}

```{r}
#| label: notes-slide-14
#| output: asis
speaker_note("slide-14")
```

## The study

```{r}
#| label: the-study-1
#| fig-width: 24
#| fig-height: 10
plot_fig_slide_15(1)
```

```{r}
#| label: notes-slide-15-1
#| output: asis
speaker_note("slide-15")
```

## The study

```{r}
#| label: the-study-2
#| fig-width: 24
#| fig-height: 10
plot_fig_slide_15(2)
```

```{r}
#| label: notes-slide-15-2
#| output: asis
speaker_note("slide-15")
```

## The study

```{r}
#| label: the-study-3
#| fig-width: 24
#| fig-height: 10
plot_fig_slide_15(3)
```

```{r}
#| label: notes-slide-15-3
#| output: asis
speaker_note("slide-15")
```

## The study

```{r}
#| label: the-study-4
#| fig-width: 24
#| fig-height: 10
plot_fig_slide_15(4)
```

```{r}
#| label: notes-slide-15-4
#| output: asis
speaker_note("slide-15")
```

## Reading the table

```{r}
#| label: reading-the-table-1
plot_fig_slide_16(1)
```

```{r}
#| label: notes-slide-16-1
#| output: asis
speaker_note("slide-16")
```

## Reading the table

```{r}
#| label: reading-the-table-2
plot_fig_slide_16(2)
```

```{r}
#| label: notes-slide-16-2
#| output: asis
speaker_note("slide-16")
```

## Reading the table

```{r}
#| label: reading-the-table-3
plot_fig_slide_16(3)
```

```{r}
#| label: notes-slide-16-3
#| output: asis
speaker_note("slide-16")
```

## Reading the table

```{r}
#| label: reading-the-table-4
plot_fig_slide_16(4)
```

```{r}
#| label: notes-slide-16-4
#| output: asis
speaker_note("slide-16")
```

## Reading the table

```{r}
#| label: reading-the-table-5
plot_fig_slide_16(5)
```

```{r}
#| label: notes-slide-16-5
#| output: asis
speaker_note("slide-16")
```

## Reading the table

```{r}
#| label: reading-the-table-6
plot_fig_slide_16(6)
```

```{r}
#| label: notes-slide-16-6
#| output: asis
speaker_note("slide-16")
```

## Reading the table

```{r}
#| label: reading-the-table-7
plot_fig_slide_16(7)
```

```{r}
#| label: notes-slide-16-7
#| output: asis
speaker_note("slide-16")
```

## Reading the table

```{r}
#| label: reading-the-table-8
plot_fig_slide_16(8)
```

```{r}
#| label: notes-slide-16-8
#| output: asis
speaker_note("slide-16")
```

## Reading the table

```{r}
#| label: reading-the-table-9
f_forest()
```

```{r}
#| label: notes-slide-16-9
#| output: asis
speaker_note("slide-16")
```

## Thank you

```{r}
#| label: notes-slide-17
#| output: asis
speaker_note("slide-17")
```
